---
title: "inspect_model_differences_in_terms_of_drift_rate"
author: "BBZ"
date: "11/13/2017"
output: html_document
---

# Overview
In this notebook, I investigate how the various parameterizations tested influence the drift rate, and the difference in drift rate across framing conditions.

# Preliminaries

Clear everything
```{r Provide access to pipe from magrittr }
rm(list = ls())
```

Make `magrittr`'s pipe accessible from notebook.
```{r Provide access to pipe from magrittr }
"%>%" <- magrittr::`%>%`
```

Function for plotting
```{r Function for plotting v}

plot_v_heatmap <- function(data) {
    ggplot2::ggplot(data = data,
                  ggplot2::aes(x = delta_m,
                               y = delta_t)
                  ) + 
    ggplot2::facet_wrap("condition") + 
    ggplot2::geom_tile(ggplot2::aes(fill = v)) + 
    ggplot2::geom_text(ggplot2::aes(label = round(v, 1)),
                       size = 2) +
    ggplot2::scale_fill_gradient2(low = "#67a9cf", mid = '#f7f7f7' , high = "#ef8a62") +
    
    ggplot2::coord_equal() + 
    
    ggplot2::ggtitle("Drift rate as a function of condition (panels), money difference (x), and time difference (y)") +
    ggplot2::scale_x_discrete(name = "Money difference between options (in Euros)") +
    ggplot2::scale_y_discrete(name = "Time difference between options (in days)") +
    
    ggplot2::theme_minimal()
}


```

Make stimuli
```{r Make stimuli}

frames <- c('delay','date')
m_l <- c(50)
pct_m_l <- seq(from = 0.10, to = 0.90, by = 0.10)
t_s <- c(0)
interval <- seq(from = 10, to = 90, by = 10)

date_delay_stimuli <- 
  itchmodel::make_stimulus_df(frames = frames,
                              m_l = m_l,
                              pct_m_l = pct_m_l,
                              t_s = t_s,
                              interval = interval,
                              n_reps = 1
                              )
```

# Simulate data

## Date delay value sensitivity model

```{r Simulate data for date_delay_value_sensitivity model}

# Determine parameters
par_delay_value_sensitivity <- 
  list('alpha' = 0.85,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

par_date_value_sensitivity <- 
  list('alpha' = 0.95,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

# Make tidy data frame
tb_date_delay_value_sensitivity <- 
  stimuli %>%
  dplyr::mutate(delta_m = as.factor(m_l - m_s),
                delta_t = as.factor(t_l - t_s),
                v_delay = compute_drift_rate(parameters = par_delay_value_sensitivity, 
                                             stimuli = stimuli, 
                                             parameterization = "date_delay_value_sensitivity", 
                                             frame = "delay"),
                v_date = compute_drift_rate(parameters = par_date_value_sensitivity, 
                                            stimuli = stimuli, 
                                            parameterization = "date_delay_value_sensitivity", 
                                            frame = "date"),
                v_date_minus_delay = v_date - v_delay
                ) %>%
  tidyr::gather(key = 'condition', value = 'v', v_delay:v_date_minus_delay) %>%
  dplyr::mutate(condition = factor(condition, levels = c("v_delay", "v_date", "v_date_minus_delay")))

```



## Date-delay time scaling model

```{r Simulate data for date_delay_time_scaling model}

# Determine parameters
par_delay_time_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1.1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

par_date_time_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

# Make tidy data frame
tb_date_delay_time_scaling <- 
  stimuli %>%
  dplyr::mutate(delta_m = as.factor(m_l - m_s),
                delta_t = as.factor(t_l - t_s),
                v_delay = compute_drift_rate(parameters = par_delay_time_scaling, 
                                             stimuli = stimuli, 
                                             parameterization = "date_delay_time_scaling", 
                                             frame = "delay"),
                v_date = compute_drift_rate(parameters = par_date_time_scaling, 
                                            stimuli = stimuli, 
                                            parameterization = "date_delay_time_scaling", 
                                            frame = "date"),
                v_date_minus_delay = v_date - v_delay
                ) %>%
  tidyr::gather(key = 'condition', value = 'v', v_delay:v_date_minus_delay) %>%
  dplyr::mutate(condition = factor(condition, levels = c("v_delay", "v_date", "v_date_minus_delay")))

```

## Defer-speedup effect: value scaling model

```{r Simulate defer-speedup effect with value scaling model}

# Determine parameters
par_defer_value_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1.2,
       'beta' = 0.70,
       'kappa' = 1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

par_speedup_value_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1.2,
       'beta' = 0.70,
       'kappa' = 1,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

# Make tidy data frame
tb_defer_speedup_value_scaling <- 
  stimuli %>%
  dplyr::mutate(delta_m = as.factor(m_l - m_s),
                delta_t = as.factor(t_l - t_s),
                v_defer = compute_drift_rate(parameters = par_defer_value_scaling, 
                                             stimuli = stimuli, 
                                             parameterization = "defer_speedup_value_scaling", 
                                             frame = "defer"),
                v_speedup = compute_drift_rate(parameters = par_speedup_value_scaling, 
                                            stimuli = stimuli, 
                                            parameterization = "defer_speedup_value_scaling", 
                                            frame = "speedup"),
                v_defer_minus_speedup = v_defer - v_speedup
                ) %>%
  tidyr::gather(key = 'condition', value = 'v', v_defer:v_defer_minus_speedup) %>%
  dplyr::mutate(condition = factor(condition, levels = c("v_defer", "v_speedup", "v_defer_minus_speedup")))

```

## Defer-speedup effect: time scaling model

```{r Simulate defer-speedup effect with value scaling model}

# Determine parameters
par_defer_time_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1.2,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

par_speedup_time_scaling <- 
  list('alpha' = 0.90,
       'mu' = 1,
       'beta' = 0.70,
       'kappa' = 1/1.2,
       'w' = 0.5,
       "a" = 1,
       "t0" = 1 
       )

# Make tidy data frame
tb_defer_speedup_time_scaling <- 
  stimuli %>%
  dplyr::mutate(delta_m = as.factor(m_l - m_s),
                delta_t = as.factor(t_l - t_s),
                v_defer = compute_drift_rate(parameters = par_defer_time_scaling, 
                                             stimuli = stimuli, 
                                             parameterization = "defer_speedup_time_scaling", 
                                             frame = "defer"),
                v_speedup = compute_drift_rate(parameters = par_speedup_time_scaling, 
                                            stimuli = stimuli, 
                                            parameterization = "defer_speedup_time_scaling", 
                                            frame = "speedup"),
                v_defer_minus_speedup = v_defer - v_speedup
                ) %>%
  tidyr::gather(key = 'condition', value = 'v', v_defer:v_defer_minus_speedup) %>%
  dplyr::mutate(condition = factor(condition, levels = c("v_defer", "v_speedup", "v_defer_minus_speedup")))

```

# Plot data


```{r}
plot_v_heatmap(data = tb_date_delay_value_sensitivity)
```



```{r}
plot_v_heatmap(data = tb_date_delay_time_scaling)
```


```{r}
plot_v_heatmap(data = tb_defer_speedup_value_scaling)
```


```{r}
plot_v_heatmap(data = tb_defer_speedup_time_scaling)
```

