---
title: "Parameter and model recovery"
author: "Bram B. Zandbelt"
params:
  n_repetition: 2
  i_dataset: 1
  parameterization: "defer_speedup_time_scaling"
output: 
  html_document:
    toc: true
    number_sections: true
    df_print: paged
bibliography: "/Users/bramzandbelt/surfdrive/projects/BEWITCHING/itchmodel/itchmodel.bib"
biblio-style: authoryear
---

# Overview

Steps:
- Define some variables
- Load the synthetic data
- 

# Preliminaries

## Start from a well-defined state

Set state for __random number generation__. This ensures that results can be reproduced.
```{r Set state for random number generation}
set.seed(19821101 + 4 * (params$i_dataset -1) + params$n_repetition - 1)
```

## Define some general variables

Make `magrittr`'s __pipe__ accessible from notebook. Pipes enable clear expression of a sequence of multiple operations.
```{r Provide access to pipe from magrittr }
"%>%" <- magrittr::`%>%`
```

Identify the __project directory__. All paths in the code are relative to it.
```{r Define project directory}
project_dir <- rprojroot::find_root(rprojroot::has_file("DESCRIPTION"))
```

Define the __name of the notebook__ (`notebook_name`). It is used to save notebook output to a notebook-specific directory.
```{r Define notebook name}
notebook_name <- "parameter_and_model_recovery"
```

Define some output directories and verify their existence. Optimization statistics obtained in this notebook will be written to a notebook-specific directory inside `optimizations_dir`. If directories do not exist, they are created.
```{r Define and verify existence of output directories}
optimizations_dir <- file.path(project_dir,"data","optimizations")
synthetic_data_dir <- file.path(project_dir,"data","simulations", "simulate_synthetic_data_for_parameter_and_model_recovery") 
itchmodel::verify_output_dirs(base_dirs = list(optimizations_dir),
                              notebook_name = notebook_name)
```

## Define some model variables

The __name of the model__ (`model_name`). Currently, only drift diffusion models (DDM) are implemented.
```{r Define model name}
model_name <- "DDM"
```

Synthetic data were generated with four model parameterizations: the defer/speedup intertemporal choice data were generated with the `defer_speedup_time_scaling` and `defer_speedup_value_scaling` parameterizations; the date/delay intertemporal choice data were generated with the `date_delay_time_scaling` and `date_delay_value_scaling` parameterizations. We will fit each data with the parameterization that generated the synthetic data set (`generating_parameterization`) and a "counter-parameterization".

```{r Retrieve parameterization}
# parameterization is specified in the YAML header or in the params argument of rmarkdown::render
(parameterization <- params$parameterization)
```


```{r Define model parameterizations}
generating_parameterization <- parameterization

if (parameterization == "defer_speedup_time_scaling") {
  counter_parameterization <- "defer_speedup_value_scaling"
} else if (parameterization == "defer_speedup_value_scaling") {
  counter_parameterization <- "defer_speedup_time_scaling"
} else if (parameterization == "date_delay_time_scaling") {
  counter_parameterization <- "date_delay_value_scaling"
} else if (parameterization == "date_delay_value_scaling") {
  counter_parameterization <- "date_delay_time_scaling"
}

```

Synthetic data were generated with four levels of repetition: each trial (i.e. a factorial combination of `frame`, `m_s`, `m_l`, `t_s`, and `t_l`) was repeated 2, 3, 4, or 5 times in the experiment (`n_repetition`). Here, we fit the dataset containing `n_repetition` repetitions.

```{r Retrieve number of repetitions}
# n_repetition is specified in the YAML header or in the params argument of rmarkdown::render
(n_repetition <- params$n_repetition)
```

A total of 100 synthetic data sets were generated for each combination of model, parameterization, and number of repetitions. Here, we fit the dataset with index `i_dataset`.

```{r Retrieve dataset index}
# i_dataset is specified in the YAML header or in the params argument of rmarkdown::render
(i_dataset <- params$i_dataset)
```

## Define some optimization variables

The __maximum number of iterations__ (`max_iter`) of the optimization algorithm before it should terminate.
```{r Set maximum number of iterations}
max_iter <- 1000
```

The relative tolerance (`rel_tol`) is the tolerance for the stopping criterion of the optimization algorithm.
```{r Set telative tolerance}
rel_tol <- 1e-6
```

The __number of populations per free parameter__ defined the number of candidate solutions in the randomly distributed initial population. Higher numbers increase the likelihood of converging to a global optimum.
```{r Set number of populations per free parameters}
n_pop_per_free_param <- 20
```

# Load data 

## Load the synthetic data

Define file name of __synthetic data__ file
```{r Define file name of synthetic data file}
fmt_data <- "expt_stimuli_observations_model_%s_parameterization_%s_ix_%d_nrep_%d.csv"
(synthetic_data_file <- 
  sprintf(fmt = fmt_data, 
          model_name, 
          generating_parameterization, 
          i_dataset, 
          n_repetition)
  )
```

Load synthetic data file
```{r Load synthetic data file}
(synthetic_data <- 
  readr::read_csv(file = file.path(synthetic_data_dir, synthetic_data_file),
                  col_types = readr::cols()
                  ))
```

Count the __number of data points__
```{r Count the number of data points}
(n_data_points <- nrow(synthetic_data))
```

Put synthetic data in correct format for model fitting
```{r Put synthetic data in correct format for model fitting}
synthetic_data_grpd <- 
  synthetic_data %>%
  dplyr::group_by(frame)

stimuli <- 
  synthetic_data_grpd %>%
  dplyr::select(frame, m_s, t_s, m_l, t_l, m_ss_type) %>%
  tidyr::nest()

observations <- 
  synthetic_data_grpd %>%
  dplyr::select(frame, rt, response) %>%
  tidyr::nest()

(synthetic_data_formatted <- 
dplyr::left_join(x = stimuli,
                 y = observations,
                 by = c("frame")
                 ) %>%
  dplyr::rename(stimuli = data.x,
                observations = data.y))
```

## Load the data generating model parameter values

Define the name of the file containing __data generating model parameter values__
```{r Define the name of the file containing data generating model parameter values}
fmt_params <- "model_parameters_model_%s_parameterization_%s_ix_%d.csv"
(synthetic_params_file <- 
  sprintf(fmt = fmt_params, 
          model_name, 
          generating_parameterization, 
          i_dataset))
```

Load parameter file
```{r Load parameter file}
(generating_params <- 
  readr::read_csv(file = file.path(synthetic_data_dir, synthetic_params_file),
                  col_types = readr::cols()
                  ))
```

_Parameter definitions_:  
_alpha_: sensitivity parameter of the function transforming objective into subjective value  
_mu_: scaling parameter of the function transforming objective into subjective value  
_beta_: sensitivity parameter of the function transforming objective into subjective time  
_kappa_: scaling parameter of the function transforming objective into subjective time  
_w_: parameter that weighs and scale the differences in subjective value and time, before comparison  
_a_: threshold separation of the diffusion process  
_t0_: non-decision time  

Indices following the parameter name indicate that the parameter varies across conditions.

# Fit data generating model to the synthetic data
The dataset will be fit with the `generating_parameterization`: `r generating_parameterization`.

Define parameter __bounds__
```{r Define parameter bounds of the generating_parameterization}
lowers_gen <- itchmodel::get_par_bounds(model = model_name,
                                        parameterization = generating_parameterization,
                                        bound = 'lower')
uppers_gen <- itchmodel::get_par_bounds(model = model_name,
                                        parameterization = generating_parameterization,
                                        bound = 'upper')

```

The __number of free parameters__ (`n_free_params`) can be determined by counting the number of parameters for which the lower and upper bounds are identical.
```{r Determine the number of free parameters of the generating_parameterization}
n_free_params_gen <- sum(!(lowers_gen == uppers_gen))
```

```{r Fit data generating model to synthetic data}
optim_out_gen_model <- 
  DEoptimR::JDEoptim(fn = itchmodel::get_log_likelihood, # optimization function
                     lower = lowers_gen,
                     upper = uppers_gen,
                     constr = itchmodel::get_nonlinear_constraints,
                     maxiter = max_iter,
                     tol = rel_tol,
                     NP = n_pop_per_free_param * n_free_params_gen,
                     trace = TRUE,
                     triter = 25, # Print every 25th iteration 
                     # Additional arguments passed to fn and constr:
                     data = synthetic_data_formatted,
                     model = model_name,
                     parameterization = generating_parameterization
                   )
```


```{r Print the best-fitting parameters of the generating_parameterization model fit}
(best_fit_gen_model <- 
   itchmodel::params_to_tibble(optim_out_gen_model$par, 
                               model = "DDM", 
                               parameterization = generating_parameterization)
 )
```

Print optimization procedure stats
```{r Print optimization procedure stats of the generating_parameterization model fit}
(optim_stats_gen_model <- 
   itchmodel::get_fit_stats(optim_output = optim_out_gen_model, 
                            model = model_name, 
                            parameterization = generating_parameterization, 
                            n_data_points = n_data_points)
 )
```

Plot the observations and model predictions, using best-fitting parameters
```{r Plot the observations and model predictions of the generating_parameterization, using best-fitting parameters}
(plt_generating_parameterization <- 
  itchmodel::plot_model_fit_to_data(df = synthetic_data_formatted,
                                    params = optim_out_gen_model$par,
                                    model = model_name,
                                    parameterization = generating_parameterization
                                    ))
```
Points represent observed probabilities of choosing large-but-later option. Lines represent best-fitting model predictions.

# Fit counter parameterization to synthetic data
The dataset will be fit with the `counter_parameterization`: `r counter_parameterization`.

Parameter __lower bounds__ (`lowers`) and __upper bounds__ (`uppers`)
```{r Define parameter bounds of the counter_parameterization}
lowers_counter <- itchmodel::get_par_bounds(model = model_name,
                                            parameterization = counter_parameterization,
                                            bound = 'lower')
uppers_counter <- itchmodel::get_par_bounds(model = model_name,
                                            parameterization = counter_parameterization,
                                            bound = 'upper')  
```

The __number of free parameters__ (`n_free_params`) can be determined by counting the number of parameters for which the lower and upper bounds are identical.
```{r Determine the number of free parameters of the counter_parameterization}
n_free_params_counter <- sum(!(lowers_counter == uppers_counter))
```

```{r Fit counter_parameterization model to synthetic data}
optim_out_counter_model <- 
  DEoptimR::JDEoptim(fn = itchmodel::get_log_likelihood, # optimization function
                     lower = lowers_counter,
                     upper = uppers_counter,
                     constr = itchmodel::get_nonlinear_constraints,
                     maxiter = max_iter,
                     tol = rel_tol,
                     NP = n_pop_per_free_param * n_free_params_counter,
                     trace = TRUE,
                     triter = 25, # Print every 25th iteration 
                     # Additional arguments passed to fn and constr:
                     data = synthetic_data_formatted,
                     model = model_name,
                     parameterization = counter_parameterization
                   )
```

```{r Print the best-fitting parameters of the counter_parameterization model fit}
(best_fit_counter_model <- 
   itchmodel::params_to_tibble(optim_out_counter_model$par, 
                               model = "DDM", 
                               parameterization = counter_parameterization)
 )
```

Print optimization procedure stats
```{r Print optimization procedure stats of the counter_parameterization model fit}
(optim_stats_counter_model <- 
   itchmodel::get_fit_stats(optim_output = optim_out_counter_model, 
                            model = model_name, 
                            parameterization = counter_parameterization, 
                            n_data_points = n_data_points)
 )
```

Plot the observations and model predictions, using best-fitting parameters
```{r Plot the observations and model predictions of the counter_parameterization, using best-fitting parameters}
(plt_counter_parameterization <- 
  itchmodel::plot_model_fit_to_data(df = synthetic_data_formatted,
                                    params = optim_out_counter_model$par,
                                    model = model_name,
                                    parameterization = counter_parameterization
                                    ))
```
Points represent observed probabilities of choosing large-but-later option. Lines represent best-fitting model predictions.

# Compare optimization function value and best-fitting parameters between models

Optimization function value. Lower values represent better fits
```{r}
tibble::tibble(model = c("generating", "counter")) %>%
  dplyr::bind_cols(dplyr::bind_rows(optim_stats_gen_model, optim_stats_counter_model))
```

Best-fitting parameters, relative to true generating parameters
```{r}
tibble::tibble(parameter_set = c("true", "best_fit_generating", "best_fit_counter")) %>%
  dplyr::bind_cols(dplyr::bind_rows(generating_params, best_fit_gen_model, best_fit_counter_model))
```

Best-fitting parameters of generating parameterization, relative to bounds
```{r Best-fitting parameters of generating parameterization, relative to bounds}
lowers_gen_tib <- 
  itchmodel::params_to_tibble(lowers_gen, 
                            model = model_name, 
                            parameterization = generating_parameterization)

uppers_gen_tib <- 
  itchmodel::params_to_tibble(uppers_gen, 
                            model = model_name, 
                            parameterization = generating_parameterization)

tibble::tibble(parameter_set = c("lower", 
                                 "best_fit_generating", 
                                 "upper", 
                                 "best_fit_minus_lower", 
                                 "upper_minus_best_fit")
               ) %>%
  dplyr::bind_cols(., dplyr::bind_rows(lowers_gen_tib, 
                                       best_fit_gen_model, 
                                       uppers_gen_tib,
                                       lowers_gen_tib - best_fit_gen_model, 
                                       uppers_gen_tib - best_fit_gen_model
                                       )
                   )
```

Best-fitting parameters of counter parameterization, relative to bounds
```{r Best-fitting parameters of counter parameterization, relative to bounds}
lowers_counter_tib <- 
  itchmodel::params_to_tibble(lowers_counter, 
                              model = model_name, 
                              parameterization = counter_parameterization)

uppers_counter_tib <- 
  itchmodel::params_to_tibble(uppers_counter, 
                              model = model_name, 
                              parameterization = counter_parameterization)

tibble::tibble(parameter_set = c("lower", 
                                 "best_fit_counter", 
                                 "upper", 
                                 "best_fit_minus_lower", 
                                 "upper_minus_best_fit")
               ) %>%
  dplyr::bind_cols(., dplyr::bind_rows(lowers_counter_tib, 
                                       best_fit_counter_model, 
                                       uppers_counter_tib,
                                       lowers_counter_tib - best_fit_counter_model, 
                                       uppers_counter_tib - best_fit_counter_model
                                       )
                   )
```

# Write optimization statistics to disk

## Best-fitting parameters

Write best-fitting parameters of generating parameterization to disk
```{r Write best-fitting parameters of generating parameterization to disk}

fmt_bf_params_generating_file <- "best_fitting_params_generating_model_%s_parameterization_%s_ix_%d_nrep_%d.csv"

(bf_params_generating_file <- 
  sprintf(fmt = fmt_bf_params_generating_file, 
          model_name, 
          generating_parameterization, 
          i_dataset, 
          n_repetition)
  )

readr::write_csv(x = best_fit_gen_model,
                 path = file.path(optimizations_dir, 
                                  notebook_name,
                                  bf_params_generating_file),
                 append = TRUE, # enables detection of problems: each file should have only one line,
                 col_names = TRUE
                 )
```

Write best-fitting parameters of counter parameterization to disk
```{r Write best-fitting parameters of counter parameterization to disk}
fmt_bf_params_counter_file <- "best_fitting_params_counter_model_%s_parameterization_%s_ix_%d_nrep_%d.csv"

(bf_params_counter_file <- 
  sprintf(fmt = fmt_bf_params_counter_file, 
          model_name, 
          counter_parameterization, 
          i_dataset, 
          n_repetition)
  )

readr::write_csv(x = best_fit_counter_model,
                 path = file.path(optimizations_dir, 
                                  notebook_name,
                                  bf_params_counter_file),
                 append = TRUE, # enables detection of problems: each file should have only one line,
                 col_names = TRUE
                 )
```

## Optimization statistics

Write optimization statistics of generating parameterization to disk
```{r Write optimization statistics of generating parameterization to disk}

fmt_optim_stats_generating_file <- "optimization_statistics_generating_model_%s_parameterization_%s_ix_%d_nrep_%d.csv"

(optim_stats_generating_file <- 
  sprintf(fmt = fmt_optim_stats_generating_file, 
          model_name, 
          generating_parameterization, 
          i_dataset, 
          n_repetition)
  )

readr::write_csv(x = optim_stats_gen_model,
                 path = file.path(optimizations_dir, 
                                  notebook_name,
                                  optim_stats_generating_file),
                 append = TRUE, # enables detection of problems: each file should have only one line,
                 col_names = TRUE
                 )
```

Write optimization statistics of counter parameterization to disk
```{r Write optimization statistics of counter parameterization to disk}

fmt_optim_stats_counter_file <- "optimization_statistics_counter_model_%s_parameterization_%s_ix_%d_nrep_%d.csv"

(optim_stats_counter_file <- 
  sprintf(fmt = fmt_optim_stats_counter_file, 
          model_name, 
          generating_parameterization, 
          i_dataset, 
          n_repetition)
  )

readr::write_csv(x = optim_stats_counter_model,
                 path = file.path(optimizations_dir, 
                                  notebook_name,
                                  optim_stats_counter_file),
                 append = TRUE, # enables detection of problems: each file should have only one line,
                 col_names = TRUE
                 )
```
